Robust and Efficient CSI Compression Using CLLWCsiNet in Noisy MIMO Environments

### 2025-04-10 Revise-version.tex
unfortunately we have emphazised on multiresolution of my architecture which was not in my model

\section{Proposed Model: CLLWCsiNet }
\label{sec:proposed_model}
Building on the challenges identified in Section~\ref{sec:introduction}, we now detail our efficiency optimizations.
The proposed model, CLLWCsiNet, introduces a novel approach to compressing and refining CSI feedback for MIMO systems, particularly under the constraints of 5G and 6G networks. This model utilizes convolutional latent-space to reduce parameters while maintaining predictive NMSE in noisy channels, making it suitable for resource-constrained user equipment.

Model Architecture:
The CLLWCsiNet model is an extension of the CsiNet model with several key enhancements:
\subsection{Architecture}
\subsubsection{Convolutional Latent-Space}:
This technique significantly reduces the number of parameters compared to traditional fully connected layers. This reduction is achieved by compressing the feature maps into a lower-dimensional space using convolutional layers, thus minimizing the computational overhead (Figure \ref{fig:figure1}). 
The input CSI matrix has dimensions 2 \begin{math} \times \end{math} 32 \begin{math} \times \end{math} 32, which is passed through 3 different convolutional filters with kernel sizes of (7,7), (5,5), and (3,3). Following the convolutional factorization approach proposed in \cite{abn}, the convolution kernel is split into two parts: (1,a) and (a,1), where ‘a’ is the kernel size. By factoring the convolution kernel in this manner, the model complexity is reduced while preserving the expressive power of the original convolution operation. This factorization technique has been shown to improve the efficiency of the neural network architecture without significantly sacrificing its performance.

\subsubsection{Encoder Compression}
After the initial passing of CSI matrix through the first part of the multi-resolution filtering architecture, the CSI matrix is reshaped from its original dimensions of 2 \begin{math} \times \end{math} 32 \begin{math} \times \end{math} 32 to 64 \begin{math} \times \end{math} 1 \begin{math} \times \end{math} 32. This reshaping is done through a set of selective compression filters. For a compression ratio of 8, the design selects a different set of filters, which can be switched based on the code. The compression ratio of 8 corresponds to a codeword dimension of 8 \begin{math} \times \end{math} (1 \begin{math} \times \end{math} 32). Similarly, for a compression ratio of 4, the codeword dimension becomes 16 \begin{math} \times \end{math} (1 \begin{math} \times \end{math} 32), and so on.
As shown in Figure \ref{fig:figure2}, the filtering process starts with 64 filters and progressively reduces the number of filters to 8, in two resolution passes. For a compression ratio of 4, the number of filters starts at 64 and is reduced to 16 through the multi-resolution filtering (Table \ref{table:model_comparison}).

This multi-resolution filtering approach, along with the selective compression based on the target ratio, allows the model to efficiently encode the CSI matrix while preserving the necessary information.

@@

\subsubsection{Denoising Capability}:
The model incorporates a denoising network to remove the added noise. This network uses convolutional layers to clean the CSI feedback before it is further processed, ensuring high accuracy even under noisy conditions. In a symmetric manner, the convolution filters are also present in the decoder part of the architecture. Before the multi-resolution RefineNet, we reshape the CSI matrix to 2×32×32 (Figure \ref{fig:figure3}).

\subsubsection{Convolution Factorization} To prevent an increase in FLOPs while simultaneously enhancing accuracy, we utilize convolution factorization as introduced in CRNet\cite{abn}. Additionally, we employ different resolutions to make learning the random structure of CSI more effective. Large kernel sizes, as suggested in CsiNetPlus\cite{abo}, are used to avoid futile coefficients and ensure coverage of the CSI that contains data. Conversely, CRNet\cite{abn} suggests smaller kernel sizes for learning granular features in the CSI matrix. We integrate both approaches into a comprehensive resolution block.

@@

Beyond architectural choices, we further optimize deployment through:

\subsection{Model Quantization}

Quantization is a crucial technique for optimizing deep neural networks (DNNs) to be deployed on resource-constrained devices, such as mobile phones and IoT devices. Each quantization approach offers distinct advantages and trade-offs. Neural network binarization is an extreme form of quantization where weights and activations are limited to binary values, typically -1 and 1, resulting in significant model size reduction and computational efficiency. 
\begin{equation}
	\omega_{b}=\mathrm{sign}(\omega)
\end{equation}
where \begin{math}
	\omega
\end{math}
are the original weights and \begin{math} \omega_{b}\end{math} are the binarized weights. This drastic reduction in precision requires specialized training methods to maintain accuracy. \cite{aaa1} Lu et al. \cite{abp} were pioneers in applying binary neural networks, and they later introduced vector quantization in the ACRNet\cite{abx} architecture. Their approach combined the binarization of fully connected layers with subsequent vector quantization to generate bit streams of 0s and 1s between the encoder and decoder. Convolutional quantization focuses on reducing the precision of weights and activations specifically in convolutional layers, which are prevalent in CNNs used for image processing. 
\begin{equation}
	\chi_{q}=\mathrm{round}(\frac{\chi}{S})+Z
\end{equation}
where \begin{math}
	S
\end{math} is the scale factor and \begin{math}
Z
\end{math} is the zero-point \cite{convolutionalquantization}. 
Static quantization, or post-training quantization, is performed on a pre-trained model using calibration data to determine scaling factors for weights and activations. This method reduces the precision of these elements, typically to 8-bit integers, and is suitable for deployment on fixed-resource devices with consistent input data\cite{staticquantization}.
Dynamic quantization involves quantizing weights during training and performing quantization of activations dynamically at inference time, allowing the model to adapt to varying input data distributions. This quantization process ensures that the model maintains flexibility and efficiency, adapting to different data patterns while reducing computational overhead. \cite{dynamicquantization}

@@

\begin{table*}[ht]
	\centering
	\caption{NMSE (dB) 8-bit Dynamic Quantization of Proposed Auto-Encoders for Indoor and Outdoor Scenario}
	\label{table:dynamicquantization}
	\begin{threeparttable}
		\begin{tabular}{clcccccc}
			\toprule
			\textbf{CR} & \textbf{Architecture} & \multicolumn{2}{c}{\textbf{Indoor}} & \multicolumn{2}{c}{\textbf{Outdoor}} \\ 
			& & Params & NMSE & Params & NMSE \\ 
			\midrule
			\multirow{5}{*}{4} & BACRNet1x\cite{abx} & 71K & -14.20 & 71K & -7.03 \\  
			& QACRNet1x & 2632 & -16.19 & 2632 & -9.44 \\  
			& BCsiNet\cite{abp} & 1088K & -17.25 & 1088K & -8.35 \\ 
			& BACRNet10x\cite{abx} & 91K & \textbf{-17.27} & 91K & -8.78 \\ 
			& QACRNet10x & 23K & -13.94 & 23K & \textbf{-12.23} \\ 
			\midrule
			\multirow{5}{*}{8} & BACRNet1x\cite{abx} & 38K & -11.52 & 38K & -5.52 \\  
			& QACRNet1x & 2632 & -11.33 & 2632 & -7.20 \\  
			& BCsiNet\cite{abp} & 547K & -12.39 & 547K & -6.26 \\ 
			& BACRNet10x\cite{abx} & 58K & -14.96 & 58K & -6.63 \\ 
			& QACRNet10x & 23K & \textbf{-16.66} & 23K & \textbf{-8.57} \\ 
			\midrule
			\multirow{5}{*}{16} & BACRNet1x\cite{abx} & 7.12K & -2.19 & 21K & -2.92 \\  
			& QACRNet1x & 2632 & -9.07 & 2632 & -4.67 \\  
			& BCsiNet\cite{abp} & 276K & -8.99 & 276K & -4.11 \\ 
			& BACRNet10x\cite{abx} & 42K & -11.7 & 42K & -4.63 \\ 
			& QACRNet10x & 23K & \textbf{-12.19} & 23K & \textbf{-5.75} \\ 
			\bottomrule

		\end{tabular}
		\begin{tablenotes}
			\item[*] Q Indicates dynamic quantization based architecture 
		\end{tablenotes}
	\end{threeparttable}
\end{table*}

Quantization-aware training (QAT) and post-training quantization (PTQ) are two primary approaches to implement quantization. QAT integrates the quantization process into the training phase, allowing the model to learn and compensate for the quantization errors, which results in higher accuracy for the quantized model. This involves simulating quantization during both forward and backward passes of training. PTQ, however, applies quantization to a pre-trained model, often followed by techniques such as weight clustering and fine-tuning to mitigate the loss in accuracy.

Dynamic quantization is a valuable technique for applications requiring efficient use of hardware resources and fast real-time inference, such as in wireless communications. This paper introduces dynamic quantization for the first time in this context.

% Your content here

\begin{figure*}[!t] % Use [!t] to attempt to place the figure at the top of the page
	\centering
	\begin{minipage}[b]{0.45\textwidth} % Adjust the width as needed
		\centering
		\includegraphics[width=\textwidth]{NMSEvsNofparametters.pdf}
		\caption{NMSE vs Number of Parameters}
		\label{fig:nmse-params}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.4\textwidth} % Adjust the width as needed
		\centering
		\includegraphics[width=\textwidth]{NMSEvsFLOPs.pdf}
		\caption{NMSE vs FLOPs}
		\label{fig:nmse-flops}
	\end{minipage}
\end{figure*}
% More content here

\begin{table}[htb]
	\centering
	\caption{Methods Parameters}
	\label{table:methods_params}

	\begin{tabular}{ c|cc}
		\hline
		\multicolumn{1}{c|}{Methods}      & \multicolumn{2}{c}{Params}     \\

		& \multicolumn{1}{c}{UE Params }             & \multicolumn{1}{c}{UE FLOP} \\
		\hline
		CsiNet     & 1052K      & 1094k \\
		CRNet        & 1049K      & 1235K \\
		ACRNet-1x    & 1049K      & 1235K \\
		CsiNetPlus   & 1049K      & 1462K \\
		TransNet     & 2381K      & 1054K \\

	\end{tabular}

\end{table}

@@

\begin{table}[ht]
	\centering
	\begin{threeparttable} % <-- Required wrapper
	\caption{Theoretical vs. Practical BER Comparison (SNR = 20 dB)}
	\label{tab:ber_analysis}
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline
	\textbf{Mod.} & \textbf{CR} & \textbf{NMSE (dB)} & $\sigma^2_{\text{total}}$ & \textbf{SNR$_{\text{eff}}$ (dB)} & \textbf{BER} \\
	\hline
	BPSK & 4 & -15.09 & 0.031 & 18.7 & $2.1 \times 10^{-6}$ \\
	BPSK & 8 & -11.11 & 0.089 & 15.1 & $1.5 \times 10^{-4}$ \\
	\hline
	\end{tabular}
	\begin{tablenotes} % <-- Notes section
	\item[1] $\sigma^2_{\text{est}} = 0.01$ assumes pilot SNR = 20 dB
	\item[2] NMSE values from Table~2
	\item[3] BER calculated via $Q(\sqrt{2\cdot\text{SNR}_{\text{eff}}})$
	\end{tablenotes}
	\end{threeparttable}
	\end{table}

% Figure 1
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.6\linewidth]{Model_a.pdf}
	\caption{Multiresolution Encoder}
	\label{fig:figure1}
\end{figure}

% Figure 2
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Model_b.pdf}
	\caption{Convolutional Laten-Space}
	\label{fig:figure2}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Model_c.pdf}
	\caption{Denoising Module and Multiresolution RefineNet}
	\label{fig:figure3}
\end{figure}

>>>>

\section{Proposed Model: CLLWCsiNet}
\label{sec:proposed_model}
Building on the challenges identified in Section~\ref{sec:introduction}, we now detail our efficiency optimizations.
The proposed model, CLLWCsiNet, introduces a novel approach to compressing and refining CSI feedback for MIMO systems, particularly under the constraints of 5G and 6G networks. This model utilizes convolutional latent-space to reduce parameters while maintaining predictive NMSE in noisy channels, making it suitable for resource-constrained user equipment.

\subsection{Architecture}
\subsubsection{Convolutional Latent-Space}:
This technique significantly reduces the number of parameters compared to traditional fully connected layers. This reduction is achieved by compressing the feature maps into a lower-dimensional space using convolutional layers, thus minimizing the computational overhead (Figure \ref{fig:figure2}). 
The input CSI matrix has dimensions 2 \begin{math} \times \end{math} 32 \begin{math} \times \end{math} 32, which is passed through 3 different convolutional filters with kernel sizes of (7,7), (5,5), and (3,3). Following the convolutional factorization approach proposed in \cite{abn}, the convolution kernel is split into two parts: (1,a) and (a,1), where ‘a’ is the kernel size. By factoring the convolution kernel in this manner, the model complexity is reduced while preserving the expressive power of the original convolution operation. This factorization technique has been shown to improve the efficiency of the neural network architecture without significantly sacrificing its performance.

\subsubsection{Encoder Compression}
The CSI matrix is reshaped from its original dimensions of 2 \begin{math} \times \end{math} 32 \begin{math} \times \end{math} 32 to 64 \begin{math} \times \end{math} 1 \begin{math} \times \end{math} 32. This reshaping is done through a set of selective compression filters. For a compression ratio of 8, the design selects a different set of filters, which can be switched based on the code. The compression ratio of 8 corresponds to a codeword dimension of 8 \begin{math} \times \end{math} (1 \begin{math} \times \end{math} 32). Similarly, for a compression ratio of 4, the codeword dimension becomes 16 \begin{math} \times \end{math} (1 \begin{math} \times \end{math} 32), and so on.
As shown in Figure \ref{fig:figure2}, the filtering process starts with 64 filters and progressively reduces the number of filters to 8. For a compression ratio of 4, the number of filters starts at 64 and is reduced to 16 (Table \ref{table:model_comparison}).

@@

\subsubsection{Denoising Capability}:
The model incorporates a denoising network to remove the added noise. This network uses convolutional layers to clean the CSI feedback before it is further processed, ensuring high accuracy even under noisy conditions. In a symmetric manner, the convolution filters are also present in the decoder part of the architecture. Before the refinement step, we reshape the CSI matrix to 2×32×32 (Figure \ref{fig:figure3}).

\subsubsection{Convolution Factorization} To prevent an increase in FLOPs while simultaneously enhancing accuracy, we utilize convolution factorization as introduced in CRNet\cite{abn}. We employ different kernel sizes to make learning the random structure of CSI more effective. Large kernel sizes, as suggested in CsiNetPlus\cite{abo}, are used to avoid futile coefficients and ensure coverage of the CSI that contains data. Conversely, CRNet\cite{abn} suggests smaller kernel sizes for learning granular features in the CSI matrix. We integrate both approaches into a comprehensive block.

@@

\subsection{Model Quantization}
Quantization is a crucial technique for optimizing deep neural networks (DNNs) to be deployed on resource-constrained devices, such as mobile phones and IoT devices. Each quantization approach offers distinct advantages and trade-offs. Neural network binarization is an extreme form of quantization where weights and activations are limited to binary values, typically -1 and 1, resulting in significant model size reduction and computational efficiency. 
\begin{equation}
	\omega_{b}=\mathrm{sign}(\omega)
\end{equation}
where \begin{math}
	\omega
\end{math}
are the original weights and \begin{math} \omega_{b}\end{math} are the binarized weights. This drastic reduction in precision requires specialized training methods to maintain accuracy. \cite{aaa1} Lu et al. \cite{abp} were pioneers in applying binary neural networks, and they later introduced vector quantization in the ACRNet\cite{abx} architecture. Their approach combined the binarization of fully connected layers with subsequent vector quantization to generate bit streams of 0s and 1s between the encoder and decoder. Convolutional quantization focuses on reducing the precision of weights and activations specifically in convolutional layers, which are prevalent in CNNs used for image processing. 
\begin{equation}
	\chi_{q}=\mathrm{round}(\frac{\chi}{S})+Z
\end{equation}
where \begin{math}
	S
\end{math} is the scale factor and \begin{math}
Z
\end{math} is the zero-point \cite{convolutionalquantization}. 
Static quantization, or post-training quantization, is performed on a pre-trained model using calibration data to determine scaling factors for weights and activations. This method reduces the precision of these elements, typically to 8-bit integers, and is suitable for deployment on fixed-resource devices with consistent input data\cite{staticquantization}.
Dynamic quantization involves quantizing weights during training and performing quantization of activations dynamically at inference time, allowing the model to adapt to varying input data distributions. This quantization process ensures that the model maintains flexibility and efficiency, adapting to different data patterns while reducing computational overhead. \cite{dynamicquantization}

@@

\begin{figure*}[!t]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NMSEvsNofparametters.pdf}
		\caption{NMSE vs Number of Parameters}
		\label{fig:nmse-params}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.4\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NMSEvsFLOPs.pdf}
		\caption{NMSE vs FLOPs}
		\label{fig:nmse-flops}
	\end{minipage}
\end{figure*}

\begin{table*}[ht]
	\centering
	\caption{NMSE (dB) 8-bit Dynamic Quantization of Proposed Auto-Encoders for Indoor and Outdoor Scenario}
	\label{table:dynamicquantization}
	\begin{threeparttable}
		\begin{tabular}{clcccccc}
			\toprule
			\textbf{CR} & \textbf{Architecture} & \multicolumn{2}{c}{\textbf{Indoor}} & \multicolumn{2}{c}{\textbf{Outdoor}} \\ 
			& & Params & NMSE & Params & NMSE \\ 
			\midrule
			\multirow{5}{*}{4} & BACRNet1x\cite{abx} & 71K & -14.20 & 71K & -7.03 \\  
			& QACRNet1x & 2632 & -16.19 & 2632 & -9.44 \\  
			& BCsiNet\cite{abp} & 1088K & -17.25 & 1088K & -8.35 \\ 
			& BACRNet10x\cite{abx} & 91K & \textbf{-17.27} & 91K & -8.78 \\ 
			& QACRNet10x & 23K & -13.94 & 23K & \textbf{-12.23} \\ 
			\midrule
			\multirow{5}{*}{8} & BACRNet1x\cite{abx} & 38K & -11.52 & 38K & -5.52 \\  
			& QACRNet1x & 2632 & -11.33 & 2632 & -7.20 \\  
			& BCsiNet\cite{abp} & 547K & -12.39 & 547K & -6.26 \\ 
			& BACRNet10x\cite{abx} & 58K & -14.96 & 58K & -6.63 \\ 
			& QACRNet10x & 23K & \textbf{-16.66} & 23K & \textbf{-8.57} \\ 
			\midrule
			\multirow{5}{*}{16} & BACRNet1x\cite{abx} & 7.12K & -2.19 & 21K & -2.92 \\  
			& QACRNet1x & 2632 & -9.07 & 2632 & -4.67 \\  
			& BCsiNet\cite{abp} & 276K & -8.99 & 276K & -4.11 \\ 
			& BACRNet10x\cite{abx} & 42K & -11.7 & 42K & -4.63 \\ 
			& QACRNet10x & 23K & \textbf{-12.19} & 23K & \textbf{-5.75} \\ 
			\bottomrule
		\end{tabular}
		\begin{tablenotes}
			\item[*] Q Indicates dynamic quantization based architecture 
		\end{tablenotes}
	\end{threeparttable}
\end{table*}

\begin{table}[htb]
	\centering
	\caption{Methods Parameters}
	\label{table:methods_params}
	\begin{tabular}{ c|cc }
		\hline
		\multicolumn{1}{c|}{Methods}      & \multicolumn{2}{c}{Params}     \\
		& \multicolumn{1}{c}{UE Params}             & \multicolumn{1}{c}{UE FLOP} \\
		\hline
		CsiNet     & 1052K      & 1094k \\
		CRNet        & 1049K      & 1235K \\
		ACRNet-1x    & 1049K      & 1235K \\
		CsiNetPlus   & 1049K      & 1462K \\
		TransNet     & 2381K      & 1054K \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}[ht]
	\centering
	\begin{threeparttable}
	\caption{Theoretical vs. Practical BER Comparison (SNR = 20 dB)}
	\label{tab:ber_analysis}
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline
	\textbf{Mod.} & \textbf{CR} & \textbf{NMSE (dB)} & $\sigma^2_{\text{total}}$ & \textbf{SNR$_{\text{eff}}$ (dB)} & \textbf{BER} \\
	\hline
	BPSK & 4 & -15.09 & 0.031 & 18.7 & $2.1 \times 10^{-6}$ \\
	BPSK & 8 & -11.11 & 0.089 & 15.1 & $1.5 \times 10^{-4}$ \\
	\hline
	\end{tabular}
	\begin{tablenotes}
	\item[1] $\sigma^2_{\text{est}} = 0.01$ assumes pilot SNR = 20 dB
	\item[2] NMSE values from Table~2
	\item[3] BER calculated via $Q(\sqrt{2\cdot\text{SNR}_{\text{eff}}})$
	\end{tablenotes}
	\end{threeparttable}
\end{table}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.6\linewidth]{Model_a.pdf}
	\caption{Encoder Architecture}
	\label{fig:figure1}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Model_b.pdf}
	\caption{Convolutional Latent-Space}
	\label{fig:figure2}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Model_c.pdf}
	\caption{Denoising Module and Refinement}
	\label{fig:figure3}
\end{figure}
